---
title: "Introduction to Statistical Learning"
output: html_notebook
---

# Lab 1: Introduction to R

## Basic commands

```{r}
# creating a vector of numbers
x = c(1, 2, 3, 4)

# a sequence of number
z = seq(1, 10)

# we can also add step
z = seq(-pi, pi, length=50)

# printing
print(z)
print(x)

# accessing the value of x[i]
# index starts at 1 not like python
y = x[1]
print(y)

# length of the vector
len = length(x)
cat("The length of x is: ", len)

# basic operations
x = 3
y = 5

# addition
print(x + y)

# substraction
print(x - y)

# multiplication
print(x * y)

# division
print(x / y)

# power
print(x^2)

# listing all saved objects
list_ob = ls()
print(list_ob)

# to delete all stored objects 
rm(list = ls())

# if we try to call x an erro will occur
#print(x)

# if we want to use a function be we don't know how to use it exactly
?matrix

# let's create one
mat = matrix(data = c(1, 2, 3, 4), nrow = 2, ncol = 2, byrow = TRUE)
print(mat)

# we can access the elements through indexation
print(mat[1,1])
print(mat[1,])
print(mat[,2])

# we can also generate random vectors
# normal distribution
rand_vec = rnorm(10)
print(rand_vec)
# see also, pnorm, qnorm, dbnorm
# https://www.stat.umn.edu/geyer/old/5101/rlook.html#:~:text=rnorm%20is%20the%20R%20function,standard%20deviation%20of%20the%20distribution.

# for the sake of reproducibility, we can fix a seed
set.seed(42)
x = rnorm(n = 20, mean = 0, sd=1)
y = 7 + 2*x

# variance, correlation, mean, standard deviation
print(var(x))
print(mean(x))
print(sd(y))
print(cor(x, y))
```

## Plotting graphics

```{r}
# simplest way to plot a graphic
x = rnorm(100)
y = rnorm(100)
# scatter plot
plot(x, y)
```

```{r}
# we can make it better by adding a title and labels
plot(x, y, xlab="X-axis",
     ylab="Y-axis", 
     col="green",
     main="Title of our Plot")
```

```{r}
# these plots can also be saved
# as pdf
pdf("../plots/first_plot.pdf")
plot(x, y, xlab="X-axis",
     ylab="Y-axis",
     col="green",
     main="Title of our Plot")
# then, we tell R that we are done
dev.off()

# we can also save it as a jpeg, png
png("../plots/second_plot.png")
plot(x, y, xlab="X-axis",
     ylab="Y-axis",
     col="red",
     main="Title of our Plot")
# then, we tell R that we are done
dev.off()
```

## More sophisticated plots

```{r}
# contours
# we use it for 3D data as a topolgical map
# x, y should be in increasing order
x = seq(-pi, pi, length = 50)
y = x
f = outer(x, y, function(x, y) cos(x) + sin(y))
plot.new()
contour(x, y, f, nlevels = 45, add = T)

# with a color map
image(x, y, f)

# or 3D
persp(x, y, f)

# we can change the angle of view

persp(x, y, f,
      theta = 30,
      phi = 60)
```

## Loading data

```{r}
# we use the function read.table()
auto = read.table("../datasets/Auto.data")

# we can view it
View(auto)

# see the first rows
head(auto)
```

Sometimes, `R` may assume that the variable names are part of the data, thus we need to tell him that the first line is actually the header

```{r}
# header = T (or TRUE)
Auto = read.table("../datasets/Auto.data",
                  header = T,
                  na.strings = "?",
                  stringsAsFactors = T)
```

The `na.strings` tells `R` that the variables containing the sign `?` represent a missing element, as for `stringsAsFactors` , it tells `R` to consider variables filled with characters as **qualitative** variables.

We can also use the `csv` format datasets.

```{r}
# saving in csv format
write.csv(Auto, "../datasets/auto.csv")
```

```{r}
# reloading
auto = read.csv("../datasets/auto.csv",
                header = T,
                  na.strings = "?",
                  stringsAsFactors = T)
View(auto)

# dimension of our dataset
print(dim(auto))
```

We can also check names of our variables.

```{r}
names(auto)
```

And, perhaps we will need to delete rows containing missing data.

```{r}
# no missing data
auto = na.omit(auto)
print(dim(auto))
```

## Additional Graphical and Numerical Summaries

We may want to plot some of the variables in our data set, but `R` can't access them.

```{r}
# first solution
plot(auto$mpg, auto$horsepower)
```

```{r}
# second solution
attach(auto)
plot(cylinders, mpg)
```

We may want to convert some numerical variables into quantitative ones.

```{r}
# convert
cylinders = as.factor(cylinders)

#plot 
plot(cylinders, mpg)
```

A box plot is automatically produced if the x-axis variable is qualitative.

We can also plot histograms:

```{r}
hist(mpg,
     col='red')
# we can change number of breaks
hist(mpg,
     col = 5,
     breaks = 15)
```

If we want some fast representation of all the scatter plots, then we can use the function called `pairs`.

```{r}
pairs(auto)
# to identify the value of a variable in a scatter plot for a pair
plot(horsepower, mpg)
identify(horsepower, mpg, cylinders)
```

Finally, we may want to have a summary for all the central tendency and position stats for all variables

```{r}
summary(auto)
```
